---
title: "HC assignment 05"
author: "Sai Supriya"
date: '2022-11-28'
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hierarchial Clustering
```{r}
library(readr)
Cereals <- read_csv("Downloads/Cereals.csv")

str(Cereals)
```
## Installing & calling required packages
```{r}
library(tidyverse)## Data manipulation
library(cluster)  ## Clustering Algorithms
library(factoextra) ## Clustering Visualization
library(dendextend)  ## for comparing 2 dendograms
```


```{r}
Cereals1<- Cereals[, unlist(lapply(Cereals, is.numeric))]
Cereals1
```

Removing any missing values
```{r}
Cereals2 <- na.omit(Cereals1)
```
Scaling the data & Normalizing data

```{r}
Cereals3<- Cereals2[,1:13]
Cereals3<- scale(Cereals2, center = TRUE, scale = TRUE)
head(Cereals3)
```
##Task-1.Apply hierarchical clustering to the data using Euclidean distance to the normalized measurements. 
#Use Agnes to compare the clustering from single linkage, complete 
#linkage, average linkage, and Ward. Choose the best method

##Computing the distance matrix Dissimilarity matrix

```{r}
## Euclidean distance to the normalized measurements. 

Euclidean_Distance<- dist(Cereals3, method = "euclidean")


```
###Hierarchial clustering using complete linkage & plotting the obtained dendogram

```{r}
HC1<- hclust(Euclidean_Distance, method = "complete")

plot(HC1, cex=0.6, hang= -1)

round(HC1$height, 3)
```


## Using agnes and other to find best option
```{r}
library(cluster)
## Use Agnes to compare the clustering from  single linkage, complete linkage, average linkage, and Ward.


HC_single <- agnes(Cereals3, method="single")
HC_complete <- agnes(Cereals3, method = "complete")
HC_average <- agnes(Cereals3, method = "average")
HC_ward<- agnes(Cereals3, method = "ward")

print(HC_single$ac)


print(HC_complete$ac)

print(HC_average$ac)

print(HC_ward$ac)

```
It can be infered that the ac ( agglomerate coefficient for ward method is greater than the other two methods.)


## Using the dendrogram to find optimal no. of clusters
##Cluster partition A 
● Use the cluster centroids from A to assign each record in partition B (each record 
is assigned to the cluster with the closest centroid). 
● Assess how consistent the cluster assignments are compared to the 
assignments based on all the data

```{r}
HC_ward<- agnes(Euclidean_Distance, method = "ward")
pltree(HC_ward, cex= 0.6, hang= -1 , main= "Dendogram of agnes for ward")
```


#### Divisive clustering
```{r}
hc_4<- diana(Cereals3)
hc_4$dc
pltree(hc_4, cex=0.6, hang=-1, main="Dend of diana")

```
##"Srength of the endogram is observed to be 83%"

```{r}
dendrogram= hclust(dist(Cereals3, method = 'euclidean'), method = 'ward.D')
plot(dendrogram, main= paste('dendrogram'),
xlab= 'Height',
ylab= 'Euclidean distance')

```
The largest vertical distance is observed to have 5 clusters at the same level.

##Fitting hierarchial clustering model
```{r}

hc5= hclust(dist(Cereals3, method = 'euclidean'), method = 'ward.D2')


## Cut tree into 5 groups

sub_grp = cutree(hc5, k=5)

##Number of in each clusters
table(sub_grp)

##
plot(hc5, cex=0.6)
rect.hclust(hc5, k=5, border = 2:5,)
Bind_clus<- cbind.data.frame(cbind(Cereals3, sub_grp))
```
##Determing optimal clusters 

```{r}
library(cluster)
library(factoextra)
fviz_nbclust(Cereals3, FUN = hcut, method = "silhouette")
fviz_cluster(list(data= Cereals3, cluster= sub_grp))
```
###3)The elementary public schools would like to choose a set of cereals to include in their
#daily cafeterias. Every day a different cereal is offered, but all cereals should support a
#healthy diet. For this goal, you are requested to find a cluster of “healthy cereals.”
```{r}
Healthy_cereals<- cbind(Cereals2,sub_grp)
mean(Healthy_cereals[Healthy_cereals$sub_grp==1, "rating"])
mean(Healthy_cereals[Healthy_cereals$sub_grp==2,"rating"])
mean(Healthy_cereals[Healthy_cereals$sub_grp==3,"rating"])
mean(Healthy_cereals[Healthy_cereals$sub_grp==4,"rating"])
mean(Healthy_cereals[Healthy_cereals$sub_grp==5,"rating"])

##Therefore from the above observations it can be infered that the first cluster is given the high rating as the healthiest cereals of all the other 4.

```
